Enabling the Use of GPU Virtualization in Cloud Environments

### 2015-10-08 paper.tex
discussion os

\subsubsection{GPUs}
As it has been said before, an instance is based on a VM which is running on a real node with its own virtualized components, so it is possible that AWS could use a virtualization framework to offer GPU services to all the instances.
Although, the {\tt nvidia-smi} command informs us that the GPUs installed are NVIDIA GRID K520, we need to verify that they are non-virtualized devices so, we have executed the NVIDIA SDK test called {\tt bandwidthtest}. 
As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is an ad-hoc GPU. 
\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
\label{table:bwt}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccc}
Name &  Data Movement & Network & Bandwidth \\ \hline \hline
g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

@@

\subsection{Managing remote GPUs with OpenStack}
The main idea is to move from the original OpenStack architecture (see Subfigure~\ref{subfig:os-orig})
 to a solution where a new shared service takes part into the architecture and it 
is responsible for all related to GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.

\begin{figure}[htb]
  \centering
  \subfigure[Original in version XXXXXXXXXXXX]{
    \includegraphics[width=\linewidth]{images/os-orig.png}
    \label{subfig:os-orig}
   }
   \quad
  \subfigure[With GPGPU developed module]{
    \includegraphics[width=\linewidth]{images/os1.jpg}
    \label{subfig:os-mod}
   }
  \caption{OpenStack Architecture}
  \label{fig:os}
\end{figure}

@@

\subsection{Working Modes}
The developed module will allow the users to create any of the following scenarios (see Figure~\ref{fig2}). 
The users are given two configuration options to decide whether a physical GPU will be completely allocated for an instance (first column) or the instance will address a partition of the GPU as if it were a device (second column).
We refer to this as the ``mode'' and the possible values for this parameter are: ``exclusive'' or ``shared''. 
Let us suppose to have 4 GPU devices in the cluster (no matter where they were hosted, neither if they were in the same server). 
A well illustrated example of those scenarios can be seen in the first row of the  Figure~\ref{fig2}.
So, while in ``exclusive'' mode the instance is monopolizing all the GPUs, in the ``shared'' mode the GPUs have been partitioned. 
As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs.

@@

\subsection{User Interface}
Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
First of all, the Instance Launch Panel should be provided with a new field, 
where the user could assign an existent GPU pool, create a new one or leave the instance without accelerators of this kind.
As soon as the option ``New GPU Pool'' is chosen, the proper fields for the pool configuration appears(see Figure~\ref{fig:ui-launch}).
Furthermore, a new panel with the existent GPUs displays all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).

@@

\subsection{Discussion}
\label{sec:discussion}
%-Volver a hacer un resumen

\section{\uppercase{Conclusions}}
\label{sec:conclusions}

>>>>

\subsubsection{GPUs}
As it has been said before, an instance is based on a VM which is running on a real node with its own virtualized components, so it is possible that AWS could use a virtualization framework to offer GPU services to all the instances.
Although, the {\tt nvidia-smi} command informs us that the GPUs installed are NVIDIA GRID K520, we need to verify that they are non-virtualized devices so, we have executed the NVIDIA SDK test called {\tt bandwidthtest}. 
As it is shown in the Table~\ref{table:bwt}, as the bandwidth achieved in this test is higher than the network bandwith, it reveals that it is an ad-hoc GPU. 
% \begin{table}[htb]
% \renewcommand{\arraystretch}{1.3}
% \caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
% \label{table:bwt}
% \tabcolsep=0.09cm
% \begin{center}\begin{tabular}{cccc}
% Name &  Data Movement & Network & Bandwidth \\ \hline \hline
% g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
% g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
% g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
% g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
% \end{tabular}\end{center}\end{table}

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
\label{table:bwt}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{ccc}
Name &  Data Movement & Bandwidth \\ \hline \hline
g2.2xlarge & Host to Device & 3,004 MB/s \\ \hline
g2.2xlarge & Device to Host & 2,809 MB/s\\ \hline
g2.8xlarge & Host to Device & 2,814 MB/s\\ \hline
g2.8xlarge & Device to Host & 3,182 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

@@

\subsection{Managing remote GPUs with OpenStack}
The main idea is to move from the original OpenStack architecture (see Subfigure~\ref{subfig:os-orig})
 to a solution where a new shared service takes part into the architecture and it 
is responsible for all about GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
which splits the HTTP query in order to make use of both, the GPGPU API for GPU-related operations, and the nova API for the rest of the computation. 
The new GPGPU Service has brought a new feature which brings more flexibility, apart from the regular idea of having GPUs in a VM, from now on, we can create ``GPU-pools''. 
A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.

\begin{figure}[htb]
  \centering
  \subfigure[Version Icehouse]{
    \includegraphics[width=\linewidth]{images/os-orig.png}
    \label{subfig:os-orig}
   }
   \quad
  \subfigure[With GPGPU developed module]{
    \includegraphics[width=\linewidth]{images/os1.jpg}
    \label{subfig:os-mod}
   }
  \caption{OpenStack Architecture}
  \label{fig:os}
\end{figure}

@@

\subsection{Working Modes}
The developed module will allow the users to create any of the following scenarios (see Figure~\ref{fig2}). 
The users are given two configuration options to decide whether a physical GPU will be completely reserved for an instance (first column) or the instance will address a partition of the GPU as if it were a real device (second column).
We refer to this as the ``mode'' and the possible values for this parameter are: ``exclusive'' or ``shared''. 
Let us suppose to have 4 GPU devices in the cluster (no matter where they were hosted, neither if they were in the same server). 
A well illustrated example of those scenarios can be seen in the first row of the  Figure~\ref{fig2}.
So, while in ``exclusive'' mode the instance is monopolizing all the GPUs, in the ``shared'' mode the GPUs have been partitioned. 
As a result of halving the GPU memory of the accelerators, the instance will be able to work with up to 8 GPUs.

@@

\subsection{User Interface}
Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
First of all, the Instance Launch Panel should be provided with a new field, 
where the user could assign an existent GPU pool, create a new one or leave the instance without accelerators of this kind.
As soon as the option ``New GPU Pool'' is chosen, the proper fields for the pool configuration would appear (see Figure~\ref{fig:ui-launch}).
Furthermore, a new panel with the existent GPUs displays all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).

@@

\subsection{Discussion}
Again, the networking issue is restricting the performance of our executions, but it has been discussed during virtually all the paper.
However, it is not a new topic and in~\cite{tonithesis} is explained that investing in the network infrastructure could make the difference.

\section{\uppercase{Conclusions}}
\label{sec:conclusions}
We have presented an upgrade for Openstack which easily permits to create GPU-instances and manage the physical GPUs to make more profit of them.
